A Machine Learning Approach to MLB Catcher Framing

### 2021-12-16 Paper.Rmd
Commit final touches to paper

```{r, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}
load("Data/pitchData.Rda")
load("Model-Outputs/random-forest.Rda")

@@

```{r, echo=FALSE, fig.width=6, fig.height=2.5, warning=FALSE, message=FALSE}
library(tidyverse)
load("Model-Outputs/boosting.Rda")

@@

```{r,  fig.width=6, fig.height=3.2, echo=FALSE, message=FALSE, warning=FALSE}
load("Model-Outputs/boosting2.Rda")
final_df_gbm %>%
  group_by(trees) %>%
  summarize(accuracy = mean(accuracy)) %>%
  arrange(desc(accuracy)) %>%
  ggplot(aes(x = trees, y = accuracy)) +
  geom_line(color = "goldenrod") +
  theme_minimal() +
  labs(x = "Number of Trees",
       y = "Accuracy",
       title = "Boosting Accuracy by # of Trees")
```

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
df_rf %>%
  rename(parameter = nodesize) %>%
  bind_rows(final_df_gbm %>%
              group_by(trees) %>%
              summarize(accuracy = mean(accuracy)) %>%
              arrange(desc(accuracy)) %>%
              mutate(model = "boosting") %>%
              select(model, accuracy, trees) %>%
              rename(parameter = trees)) %>%
  arrange(desc(accuracy)) %>%
  kable
```

@@

```{r, message=FALSE,echo=FALSE, warning=FALSE}
library(gbm)
load("Final-Model/final_boosting_model.Rda")
var = summary.gbm(final_model, plotit = F)
ggplot(var, aes(x = fct_reorder(var, rel.inf), y = rel.inf, fill = rel.inf)) +
  geom_col() +
  coord_flip() +
  labs(y = "Relative Influence",
       title = "Predictor Variable Importance") +
  theme_minimal()+
  theme(axis.title.y = element_blank()) +
  guides(fill = F)
```

@@

```{r, fig.width=2, fig.height=2, cache=TRUE, echo=FALSE}
ggplot()+
  geom_path(data = sz, aes(x = x, y = y)) +
  xlim(-1.8, 1.8) +
  ylim(0.5, 5) +
  coord_equal() +
  geom_point(aes(x = .5, y = 2.5)) +
  geom_text(aes(x = 0.6, y = 2.8, label = "0.9")) +
  geom_point(aes(x = -.5, y = 4)) +
  geom_text(aes(x = -0.4, y = 4.3, label = "0.2")) +
  theme_minimal()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank()) 
```

@@

```{r, message=FALSE, echo=FALSE}
load("Data/pitchData.Rda")
# Analysis
data = data %>%
  mutate(strike_prob = predict(final_model, data, type = "response"),
         strike = as.numeric(as.character(strike)))

@@

```{r, echo=FALSE}
# Leader board
data %>%
  group_by(catcher_name) %>%
  summarize(strikes_above_avg = sum(credit),
            pitches_caught = n(),
            strikes_aa_100 = strikes_above_avg / (pitches_caught / 100)) %>%
  filter(pitches_caught > 2000) %>%
  arrange(desc(strikes_above_avg)) %>% head(10) %>%
  rename("Catcher Name" = catcher_name,
         "Strikes Above Average" = strikes_above_avg,
         "Pitches Caught" = pitches_caught,
         "Strikes Above Average Per 100 Pitches Caught" = strikes_aa_100) %>% 
  kable

>>>>

```{r, echo=FALSE, fig.width=7, fig.height=4, cache=TRUE}

load("Data/pitchData.Rda")
load("Model-Outputs/random-forest.Rda")

@@

```{r, echo=FALSE, fig.width=6, fig.height=2.5, warning=FALSE, message=FALSE}

library(tidyverse)
load("Model-Outputs/boosting.Rda")

@@

```{r,  fig.width=6, fig.height=3.2, echo=FALSE, message=FALSE, warning=FALSE}

load("Model-Outputs/boosting2.Rda")
final_df_gbm %>%
  group_by(trees) %>%
  summarize(accuracy = mean(accuracy)) %>%
  arrange(desc(accuracy)) %>%
  ggplot(aes(x = trees, y = accuracy)) +
  geom_line(color = "goldenrod") +
  theme_minimal() +
  labs(x = "Number of Trees",
       y = "Accuracy",
       title = "Boosting Accuracy by # of Trees")
```

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(knitr)
df_rf %>%
  rename(parameter = nodesize) %>%
  bind_rows(final_df_gbm %>%
              group_by(trees) %>%
              summarize(accuracy = mean(accuracy)) %>%
              arrange(desc(accuracy)) %>%
              mutate(model = "boosting") %>%
              select(model, accuracy, trees) %>%
              rename(parameter = trees)) %>%
  arrange(desc(accuracy)) %>%
  kable
```

@@

```{r, message=FALSE,echo=FALSE, warning=FALSE}

library(gbm)
load("Final-Model/final_boosting_model.Rda")
var = summary.gbm(final_model, plotit = F)
ggplot(var, aes(x = fct_reorder(var, rel.inf), y = rel.inf, fill = rel.inf)) +
  geom_col() +
  coord_flip() +
  labs(y = "Relative Influence",
       title = "Predictor Variable Importance") +
  theme_minimal()+
  theme(axis.title.y = element_blank()) +
  guides(fill = F)
```

@@

```{r, fig.width=2, fig.height=2, cache=TRUE, echo=FALSE}

ggplot()+
  geom_path(data = sz, aes(x = x, y = y)) +
  xlim(-1.8, 1.8) +
  ylim(0.5, 5) +
  coord_equal() +
  geom_point(aes(x = .5, y = 2.5)) +
  geom_text(aes(x = 0.6, y = 2.8, label = "0.9")) +
  geom_point(aes(x = -.5, y = 4)) +
  geom_text(aes(x = -0.4, y = 4.3, label = "0.2")) +
  theme_minimal()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank()) 
```

@@

```{r, message=FALSE, echo=FALSE}

load("Data/pitchData.Rda")
# Analysis
data = data %>%
  mutate(strike_prob = predict(final_model, data, type = "response"),
         strike = as.numeric(as.character(strike)))

@@

```{r, echo=FALSE}

# Leader board
data %>%
  group_by(catcher_name) %>%
  summarize(strikes_above_avg = sum(credit),
            pitches_caught = n(),
            strikes_aa_100 = strikes_above_avg / (pitches_caught / 100)) %>%
  filter(pitches_caught > 2000) %>%
  arrange(desc(strikes_above_avg)) %>% head(10) %>%
  rename("Catcher Name" = catcher_name,
         "Strikes Above Average" = strikes_above_avg,
         "Pitches Caught" = pitches_caught,
         "Strikes Above Average Per 100 Pitches Caught" = strikes_aa_100) %>% 
  kable
